Giant Low Surface Brightness Galaxies As Merger Remnants

### 2015-04-14 glsb.tex
fixed spelling

GLSBs are typically classified as spriral galaxies \citep{Das13}, however, high resolution imaging of Malin~1 has shown that the central region appears to be a normal

@@

\citet{Salim10} Looks at HST imaging of z = 0.1 early type galaxies that have UV excess, and finds some extended disks.  They think the gas and star formation is comming from minor mergers or accretion.

@@

To make a comparison sample of galaxies from SDSS, we select galaxies with redshifts between 0.01 and 0.07 and de-reddned $r$ magnitude between 13.5 and 17.7.  When comparing to just elliptical galaxies we make the additional color-cut of $u-r > 2.1$ (and some morphology selectors).  These are simply galaxies for comparison, not a rigorously defined volume-complete sample.  Our full SQL querries are shown in \S\ref{SQL}.  This gives us a sample of 141,001 objects, 45,703 of which have red colors.  Throughout, we use the CosmoloPy\footnote{\url{http://roban.github.com/CosmoloPy/}} fiducial cosmology ($h=0.704$, $\Omega_{m_0}=0.2726$, $\Omega_{\lambda_0}=0.7274$).  Note that we are not correcting for Malmquist bias or biases in the SDSS spectroscopic selection criteria.  Our selection is simply for galaxies of a similar magnitude and redshift.

@@

%/astro/store/shared-scratch1/yoachim/Research/GLSB/THINGS/thingsplot.py
\begin{figure}
\epsscale{.5}
\plotone{Plots/things.eps}
\epsscale{1}
\caption{ Comparison of the GLSB disks with other galaxy disks from the literature.  THINGS data taken from \citet{DeBlok2008}, Malin 1 from \citet{Moore06}, Malin 2 and UGC 6614 from \citet{Pickering97}.  Here, $R_{max}$ is the maximum radius with an \HI\ detection for the THINGS galaxies and the maximum redius with a stellar flux detection for the GLSB galaxies.  It is interesing to note that the stellar disks of GLSB galaxies are often larger than the \HI\ disks of nearby systems.  \label{disk_comp}}
\end{figure}

@@

XXX--We are currently observing higher spatial resolution IFU data with Gemini (GN-2015A-Q-75) of the central region of these galaxies to confim they are normal elliptical galaxies and to help constrain a more sophisticated dynamical model.

@@

A problem with fitting observed galaxy spetra with templates is that residuals between the model and data are dominated by non-Gaussian systematic effects such as template mis-match and improper parameterization of the SFH.  Since the residuals are non-Guassian, there is no statistical formalism for generating uncertanties in the fit.  To generate our error bars, we look at the range of fits which have $\chi^2$ values within 100 of the minimum.  In many cases, this eliminates nearly all fits except for the minimum.  In these cases, we use the 4 best-fitting models to estimate the range of possible fits.

@@

We can't really use the stellar luminosity of the disk to constrian the projenitor since there are signs of current {\emph in situ} star formation in the disks.  However, the very low stellar metalicities point to low-mass projenitors.

@@

These systems are rare enough, that one might imagine rather complicated scenarios for creating them.  For example, a three-body enocounter could eject one galaxy while leaving a second in a nearly circular orbit.  Back of the envelope, a galaxy ejected at 5000 \kms would go 1.5 degrees per gigayear.  Even if the ejection was partly along the line of sight, it would be hard to find several gigayears later.  Unlike the collisional ring scenario where one would expect the galaxy that stimulated the ring formation to still be nearby.

@@

We have observed two giant low surface brightness galaxies with the VIRUS-P IFU.  We find the systems are best explained as massive elliptical galaxies which have managed to accrete stars and gas into a lareg disk.  The stellar stars have a disk-like velocity field, and are dominated by old stars, but do show signs of ongoing star formation.

In a future work, we plan to combine our IFU observations with higher spatial resolution long-slit specta from the Apache Point 3.5m telescope to determine if the central ellipticals are dynamically distinct from their disks.  This data will also help up build dynamicall models for the systems.

@@

% LocalWords:  Debattista Horrocks Lancashire RCUK Malin Stanonik Spavone Lelli
% LocalWords:  Bothun GLSB Das Mapelli Weij Shmitz Loebman SungWon Kwak Scd Sd
% LocalWords:  Schiminovich Iodice Belokurov Fairchild RMS pPXF Cappellari th
% LocalWords:  GANDALF Sarzi CosmoloPy Bernardi GLSBs DeBlok Verheijen Yoachim
% LocalWords:  SEDs Brookhaven Instituto Astrofisica Canarias Notre JINA PR1 DM
% LocalWords:  2HE IFU UGC HSB S0 SAURON ellipticals SB0 M31 Malin2 Spitzer IC
% LocalWords:  kpc HIPASS CCD metallicity MANGA 0pt SDSS arcsec lensing 7m IRAF
% LocalWords:  Spectrograph spectrograph IDL longslit photometric FWHM SSP mis
% LocalWords:  dispersions Balmer decrement DR10 fiducial WMAP7 BAO H0 redshift
% LocalWords:  Faber Mg2 redshifts NGC kinemetry photometry anisotropy Gyr SFH
% LocalWords:  metallicities colliders rotators et al MNRAS de datacube Ppxf
% LocalWords:  dashed

>>>>

GLSBs are typically classified as spiral galaxies \citep{Das13}, however, high resolution imaging of Malin~1 has shown that the central region appears to be a normal

@@

\citet{Salim10} Looks at HST imaging of z = 0.1 early type galaxies that have UV excess, and finds some extended disks.  They think the gas and star formation is coming from minor mergers or accretion.

@@

To make a comparison sample of galaxies from SDSS, we select galaxies with redshifts between 0.01 and 0.07 and de-reddened $r$ magnitude between 13.5 and 17.7.  When comparing to just elliptical galaxies we make the additional color-cut of $u-r > 2.1$ (and some morphology selectors).  These are simply galaxies for comparison, not a rigorously defined volume-complete sample.  Our full SQL queries are shown in \S\ref{SQL}.  This gives us a sample of 141,001 objects, 45,703 of which have red colors.  Throughout, we use the CosmoloPy\footnote{\url{http://roban.github.com/CosmoloPy/}} fiducial cosmology ($h=0.704$, $\Omega_{m_0}=0.2726$, $\Omega_{\lambda_0}=0.7274$).  Note that we are not correcting for Malmquist bias or biases in the SDSS spectroscopic selection criteria.  Our selection is simply for galaxies of a similar magnitude and redshift.

@@

%/astro/store/shared-scratch1/yoachim/Research/GLSB/THINGS/thingsplot.py
\begin{figure}
\epsscale{.5}
\plotone{Plots/things.eps}
\epsscale{1}
\caption{ Comparison of the GLSB disks with other galaxy disks from the literature.  THINGS data taken from \citet{DeBlok2008}, Malin 1 from \citet{Moore06}, Malin 2 and UGC 6614 from \citet{Pickering97}.  Here, $R_{max}$ is the maximum radius with an \HI\ detection for the THINGS galaxies and the maximum radius with a stellar flux detection for the GLSB galaxies.  It is interesting to note that the stellar disks of GLSB galaxies are often larger than the \HI\ disks of nearby systems.  \label{disk_comp}}
\end{figure}

@@

XXX--We are currently observing higher spatial resolution IFU data with Gemini (GN-2015A-Q-75) of the central region of these galaxies to confirm they are normal elliptical galaxies and to help constrain a more sophisticated dynamical model.

@@

A problem with fitting observed galaxy spectra with templates is that residuals between the model and data are dominated by non-Gaussian systematic effects such as template mis-match and improper parameterization of the SFH.  Since the residuals are non-Gaussian, there is no statistical formalism for generating uncertainties in the fit.  To generate our error bars, we look at the range of fits which have $\chi^2$ values within 100 of the minimum.  In many cases, this eliminates nearly all fits except for the minimum.  In these cases, we use the 4 best-fitting models to estimate the range of possible fits.

@@

We can't really use the stellar luminosity of the disk to constrain the progenitor since there are signs of current {\emph in situ} star formation in the disks.  However, the very low stellar metallicities point to low-mass progenitors.

@@

These systems are rare enough, that one might imagine rather complicated scenarios for creating them.  For example, a three-body encounter could eject one galaxy while leaving a second in a nearly circular orbit.  Back of the envelope, a galaxy ejected at 5000 \kms would go 1.5 degrees per gigayear.  Even if the ejection was partly along the line of sight, it would be hard to find several gigayears later.  Unlike the collisional ring scenario where one would expect the galaxy that stimulated the ring formation to still be nearby.

@@

We have observed two giant low surface brightness galaxies with the VIRUS-P IFU.  We find the systems are best explained as massive elliptical galaxies which have managed to accrete stars and gas into a large disk.  The stellar stars have a disk-like velocity field, and are dominated by old stars, but do show signs of ongoing star formation.

In a future work, we plan to combine our IFU observations with higher spatial resolution long-slit spectra from the Apache Point 3.5m telescope to determine if the central ellipticals are dynamically distinct from their disks.  This data will also help up build dynamical models for the systems.

@@

% LocalWords:  Debattista Horrocks Lancashire RCUK Malin Stanonik Spavone Lelli
% LocalWords:  Bothun GLSB Das Mapelli Weij Shmitz Loebman SungWon Kwak Scd Sd
% LocalWords:  Schiminovich Iodice Belokurov Fairchild RMS pPXF Cappellari th
% LocalWords:  GANDALF Sarzi CosmoloPy Bernardi GLSBs DeBlok Verheijen Yoachim
% LocalWords:  SEDs Brookhaven Instituto Astrofisica Canarias Notre JINA PR1 DM
% LocalWords:  2HE IFU UGC HSB S0 SAURON ellipticals SB0 M31 Malin2 Spitzer IC
% LocalWords:  kpc HIPASS CCD metallicity MANGA 0pt SDSS arcsec lensing 7m IRAF
% LocalWords:  Spectrograph spectrograph IDL longslit photometric FWHM SSP mis
% LocalWords:  dispersions Balmer decrement DR10 fiducial WMAP7 BAO H0 redshift
% LocalWords:  Faber Mg2 redshifts NGC kinemetry photometry anisotropy Gyr SFH
% LocalWords:  metallicities colliders rotators et al MNRAS de datacube Ppxf
% LocalWords:  dashed CDM Kasparova LSB Dutton Raskutti Sa Morganti
%%  LocalWords:  Emonts Weijmans Ibata McConnachie Schweizer Magain
%%  LocalWords:  Oosterloo baryonic Buote GALEX Salim Kelson outlier
%%  LocalWords:  Bruzual Ahn Willett SQL Chilingarian Saulder Swaters
%%  LocalWords:  Krajn GN Kuntschner situ indices Emsell Capp CasJobs
